Sparse Function-space Representation \\ of Neural Networks

### 2023-05-17 paper/main.tex
added mnist results

\section{Maintaining a sparse function-space representation in sequential learning}
\label{sec:sequential}
%
We have presented a method for converting \todo{Is it really for any, no regularity at all required? Would remove that word if we're not sure} any trained neural network into a sparse function-space representation. The sparse representation opens an avenue for maintaining (and updating) a representation of the neural network, an important feature for sequential learning, such as continual learning.\looseness-1
%In this section, we demonstrate how sequential learning methods can benefit from our function-space representation and it's associated uncertainty estimates.

\paragraph{\our for continual learning}
% Definition of CL -> a sequence of tasks 
In the continual learning setting, the training is divided into $T$ tasks, each with its training data set $\dataset_t = \{(\mathbf{x}_{i}, \mathbf{y}_{i})\}_{i=1}^{N_t}$, which after the task is discarded and cannot be accessed in the following stages of the training. Rehearsal and function regularization-based models keep a subset of the training data for each task to help the model alleviate forgetting. In the same way, after each task, we can use \our method to build a compact representation of the neural network and then exploit this representation to construct a functional regularizer to add to the training loss for the subsequent tasks. Recently proposed GP-based regularizers \cite{ pan2020continual, rudner2022continual} have shown better performances than weight space equivalents getting state-of-the-art \todo{Could add a noun after state-of-the-art} on CL benchmarks. However, they are limited to using subset approximations of the posterior. Instead, we use the sparse data term posterior, see \cref{app:cl} for details.
%, which in practice means randomly selecting some data computing its posterior at only those points and then using that as the regularizer. 
\our is a more principled approach since the resultant GP is a low-rank approximation of the full data which ensures information is not lost. %, essentially what the sparse dual variables do.   \todo{this term is replacing the missing data}
% Where the regularizer comes from

In practice this means, that at the end of the training for each task $t$, we compute the approximated kernel function for the current model and randomly select $M$ inducing points. Given the issues described in the previous sections, related to computing the full covariance posterior $\MKxx$ for $\dataset_t$,\todo{this should be $\kappa(\MX_t, \MX_t)$} we can exploit the dual parameterization to efficiently encode the information from  $\dataset_t$ using the dual parameterization, as follows, 
%
\begin{equation}\textstyle
 	\bar{\MB}^{-1}_t = \MKzz^{-1} \MBeta_\vu \MKzz^{-1} \in \R^{M \times M}, 
 	\quad \text{s.t.} \quad
 	\MBeta_{\vu} =  \sum_{i \in \dataset_t} \vkzi \,\hat{\beta}_{i} \, \vkzi^{\T} ,    
 	% \quad \forall t \in [1, T]
\end{equation}
%
where $\MKzz$ and $\vkzi$ are the gram matrix and vector computed on the inducing points $\mathcal{Z}_t = \{\vz_i\}_{i=1}^M$ selected from $\dataset_t$.\
%
We keep a memory buffer $\mathcal{M} = \{(\mathcal{Z}_t, \vu_t, \bar{\MB}^{-1}_t)\}_{t=1}^T$, where $\vu_t \in \R^{M}$ is a vector with entries $u_{t, i} = f_{\vw_t^*}(\vz_i), \forall \vz_i \in \mathcal{Z}_t$, are the neural network outputs computed on the inducing points with the MAP estimate $\weights_t^*$ after task $t$. For notational convenience, we restrict ourselves here to single-output neural networks and refer the reader to \cref{sec:cl_multioutput} for details on the multi-output setting.

@@

    % THE TABLE CAN BE FORMATTED MORE LIKE THIS
    \begin{tabular}{l l C{\tblw} C{\tblw} C{\tblw} C{\tblw}}
    \toprule
    & Method & ACC~$\uparrow$ & NLPD~$\downarrow$ & ECE~$\downarrow$ & OOD-AUC~$\uparrow$  \\
    \midrule
    \multirow{2}{*}{FMNIST} 
    & MAP & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & BNN predictive & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & BNN predictive \cite{todo} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & GLM predictive & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & GP predictive & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & \our (NN) & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & \our & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\    
    \midrule
    \multirow{2}{*}{CIFAR-10} 
    & MAP & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & BNN predictive & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & BNN predictive \cite{todo} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & GLM predictive & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & GP predictive & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & \our (NN) & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & \our & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\    
    \bottomrule
    \end{tabular}

>>>>

\section{Maintaining a sparse function-space representation in sequential learning}
\label{sec:sequential}
%
We have presented a method for converting any trained neural network into a sparse function-space representation. The sparse representation opens an avenue for maintaining (and updating) a representation of the neural network, an important feature for sequential learning, such as continual learning.\looseness-1
%In this section, we demonstrate how sequential learning methods can benefit from our function-space representation and it's associated uncertainty estimates.

\paragraph{\our for continual learning}
% Definition of CL -> a sequence of tasks 
In the continual learning setting, the training is divided into $T$ tasks, each with its training data set $\dataset_t = \{(\mathbf{x}_{i}, \mathbf{y}_{i})\}_{i=1}^{N_t}$, which after the task is discarded and cannot be accessed in the following stages of the training. Rehearsal and function regularization-based models keep a subset of the training data for each task to help the model alleviate forgetting. In the same way, after each task, we can use \our method to build a compact representation of the neural network and then exploit this representation to construct a functional regularizer to add to the training loss for the subsequent tasks. Recently proposed GP-based regularizers \cite{ pan2020continual, rudner2022continual} have shown better performances than weight space equivalents getting state-of-the-art on CL benchmarks. However, they are limited to using subset approximations of the posterior. Instead, we use the sparse data term posterior see \cref{app:cl} for details.
%, which in practice means randomly selecting some data computing its posterior at only those points and then using that as the regularizer. 
\our is a more principled approach since the resultant GP is a low-rank approximation of the full data which ensures information is not lost. %, essentially what the sparse dual variables do.   \todo{this term is replacing the missing data}
% Where the regularizer comes from

In practice this means,  at the end of the training for each task $t$, we compute the approximated kernel function for the current model and randomly select $M$ inducing points. Given the issues described in the previous sections, related to computing the full covariance posterior $\MKxx$ for $\dataset_t$,\todo{this should be $\kappa(\MX_t, \MX_t)$} we can exploit the dual parameterization to efficiently encode the information from  $\dataset_t$ using the dual parameterization, as follows, 
%
\begin{equation}\textstyle
 	\bar{\MB}^{-1}_t = \MKzz^{-1} \MBeta_\vu \MKzz^{-1} \in \R^{M \times M}, 
 	\quad \text{s.t.} \quad
 	\MBeta_{\vu} =  \sum_{i \in \dataset_t} \vkzi \,\hat{\beta}_{i} \, \vkzi^{\T} ,    
 	% \quad \forall t \in [1, T]
\end{equation}
%
where $\MKzz$ and $\vkzi$ are the gram matrix and vector computed on the inducing points $\mathcal{Z}_t = \{\vz_i\}_{i=1}^M$ selected from $\dataset_t$.\
%
We keep a memory buffer $\mathcal{M} = \{(\mathcal{Z}_t, \vu_t, \bar{\MB}^{-1}_t)\}_{t=1}^T$, where $\vu_t \in \R^{M}$ is a vector with entries $u_{t, i} = f_{\vw_t^*}(\vz_i), \forall \vz_i \in \mathcal{Z}_t$, are the neural network outputs computed on the inducing points with the MAP estimate $\weights_t^*$ after task $t$. For notational convenience, we restrict ourselves here to single-output neural networks and refer the reader to \cref{sec:cl_multioutput} for details on the multi-output setting.

@@

    % THE TABLE CAN BE FORMATTED MORE LIKE THIS
    % \begin{tabular}{l l C{\tblw} C{\tblw} C{\tblw} C{\tblw}}
    % \toprule
    % & Method & ACC~$\uparrow$ & NLPD~$\downarrow$ & ECE~$\downarrow$ & OOD-AUC~$\uparrow$  \\
    % \midrule
    % \multirow{2}{*}{FMNIST}
    % & MAP & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    % & BNN predictive & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    % & BNN predictive \cite{todo} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    % & GLM predictive & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    % & GP predictive & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    % & \our (NN) & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    % & \our & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    % \midrule
    % \multirow{2}{*}{CIFAR-10}
    % & MAP & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    % & BNN predictive & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    % & BNN predictive \cite{todo} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    % & GLM predictive & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    % & GP predictive & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    % & \our (NN) & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    % & \our & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    % \bottomrule
    % \end{tabular}
  \begin{tabular}{l l C{\tblw} C{\tblw} C{\tblw} C{\tblw}}
    \toprule
    & Method & ACC~$\uparrow$ & NLPD~$\downarrow$ & ECE~$\downarrow$   \\
    \midrule
    \multirow{2}{*}{MNIST}
    & MAP & \val{98.22}{0.13} & \val{0.061}{0.004} & \val{0.006}{0.001} \\
    & BNN predictive \cite{todo} & \val{93.03}{0.13} & \val{0.369}{0.003} & \val{0.168}{0.001} \\
    & GLM predictive & \val{98.40}{0.05} & \val{0.054}{0.002} & \val{0.007}{0.001} \\
    & GP predictive $m=3200$ & \val{98.22}{0.13} & \val{0.058}{0.001} & \val{0.003}{0.000} \\
    & GP predictive NN $m=1024$ &  \val{97.70}{0.07} & \val{0.080}{0.002} & \val{0.011}{0.001} \\
    & GP predictive $m=1024$  &  \val{76.28}{1.85} & \val{0.702}{0.029} & \val{0.070}{0.022} \\
    & \our NN $m=1024$ &  \val{98.02}{0.12} & \val{0.066}{0.003} & \val{0.007}{0.001} \\
    & \our $m=1024$ &  \val{98.07}{0.06} & \val{0.063}{0.003} & \val{0.005}{0.001} \\
    \midrule
    \multirow{2}{*}{FMNIST}
    & MAP & \val{91.39}{0.11} & \val{0.258}{0.004} & \val{0.017}{0.001} \\
    & BNN predictive & \val{84.42}{0.12} & \val{0.942}{0.016} & \val{0.411}{0.008} \\
    & BNN predictive (Ritter et al.) & \val{91.20}{0.07} & \val{0.265}{0.004} & \val{0.024}{0.002} \\
    & GLM predictive & \val{92.25}{0.10} & \val{0.244}{0.003} & \val{0.012}{0.003} \\
    & GP predictive $m=3200$ & \val{91.36}{0.11} & \val{0.25}{0.004} & \val{0.007}{0.001} \\
    & GP predictive NN $m=1024$ &  \val{0}{0.07} & \val{0}{0.002} & \val{0}{0.001} \\
    & GP predictive $m=1024$  &  \val{0}{1.85} & \val{0}{0.029} & \val{0}{0.022} \\
    & \our NN $m=1024$ &  \val{0}{0.0} & \val{0}{0.003} & \val{0}{0.001} \\
    & \our $m=1024$ &  \val{0}{0.06} & \val{0}{0.003} & \val{0}{0.001} \\
    \bottomrule
  \end{tabular}

% & \sc \sc gp predictive  &  \val{76.28333}{1.849} & \val{0.70168}{0.02883} & \val{0.07049}{0.02154} \\
% & \sc \sc gp predictive NN \cite{immer2021improving} &  \val{97.70333}{0.074} & \val{0.08027}{0.00248} & \val{0.01052}{0.00097} \\
% & \sc \our &  \val{98.06667}{0.060} & \val{0.06312}{0.00284} & \val{0.00493}{0.00086} \\
% & \sc \our NN &  \val{98.02000}{0.122} & \val{0.06583}{0.00294} & \val{0.00714}{0.00096} \\
% M=512
% & \sc \our &  \val{96.05}{0.00} & \val{0.85}{0.01} & \val{0.50}{0.00} \\
% & \sc \sc gp predictive  &  \val{14.79}{0.01} & \val{2.30}{0.02} & \val{0.02}{0.01} \\
% & \sc \sc gp predictive NN \cite{immer2021improving} &  \val{91.49}{0.01} & \val{1.25}{0.01} & \val{0.61}{0.00} \\
% & \sc \our NN &  \val{96.96}{0.00} & \val{0.63}{0.00} & \val{0.40}{0.00} \\

% M=1024
